---
title: "R Notebook"
output: html_notebook
---

# 02 24 2020

## Loading the data

```{r}
library(mosaic)
```


```{r}
DelPrete <- read_excel("Delprete.xlsx")

# Changning data into a data frame

DelPrete <- as.data.frame(DelPrete)

View(DelPrete)

library(mosaic)

```

```{r}
# Subsetting the data so its just the Terry Collection

DelPreteT <- subset(DelPrete, Collection_number != 1)
DelPreteT <- subset(DelPreteT, Collection_number != 2)
DelPreteT <- subset(DelPreteT, Collection_number != 3)

```

```{r}
# Subsetting the data so its just the White Terry Collection


DelPreteTW <- subset(DelPrete, Collection_number !=1)
DelPreteTW <- subset(DelPrete, Collection_number !=2)
DelPreteTW <- subset(DelPrete, Collection_number !=3)
DelPreteTW <- subset(DelPreteTW, Ethnicity !="B")
```



## Descriptives

```{r}
favstats(DelPrete$Index_AP_inlet)
favstats(DelPrete$Index_bispinous)
favstats(DelPrete$Index_TD_outlet)
```

# Ap Inlet
```{r}

hist(DelPreteT$Index_AP_inlet)
plot(DelPreteTW$Index_AP_inlet, DelPreteTW$GM, xlab = "AP Inlet", ylab="Geometric Mean", main = "AP Inlet vs. Geometric Mean")
plot(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Inlet", ylab="Femur Head Adjusted", main ="AP Inlet vs. Femur Head")

cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715)

t.test(DelPreteT$Index_AP_inlet ~ DelPreteT$Ethnicity)

```


For the t-test, the p-value for the AP Inlet is greater than 0.05 so we cannot conclude that there is a statistical difference between the mean AP Inlet index value of the races of the Terry collection.

# TD Outlet
```{r}

hist(DelPreteT$Index_TD_outlet)
plot(DelPreteTW$Index_TD_outlet, DelPreteTW$GM, xlab = "TD Outlet", ylab="Geometric Mean", main = "TD Outlet vs. Geometric Mean")
plot(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "TD Outlet", ylab="Femur Head Adjusted", main ="TD Outlet vs. Femur Head")

cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715)

wilcox.test(DelPreteT$Index_TD_outlet ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is greater than 0.05 so we cannot conclude that the ranks for each index value are different and therefore there is no difference between the two groups.


# Bispinous
```{r}

hist(DelPreteT$Index_bispinous)
plot(DelPreteTW$Index_bispinous, DelPreteTW$GM, xlab = "Bispinous", ylab="Geometric Mean", main = "Bispsinous vs. Geometric Mean")
plot(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715, xlab = "Bispinous", ylab="Femur Head Adjusted", main ="Bispinous vs. Femur Head")

cor(DelPrete$Index_bispinous, DelPrete$femur_head_adjusted_1.715)
cor(DelPrete$Index_bispinous, DelPrete$GM)

cor.test(DelPreteTW$Index_bispinous, DelPreteTW$GM)
cor.test(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715)

wilcox.test(DelPreteT$Index_bispinous ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is greater than 0.05 so we cannot conclude that the ranks for each index value are different and therefore there is no difference between the two groups.

# Pelvic Breadth
```{r}

hist(DelPreteT$Index_pelvic_breadth)
plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$GM, xlab = "Pelvic Breadth", ylab="Geometric Mean", main = "Pelvic Breadth vs. Geometric Mean")
plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715, xlab = "Pelvic Breadth", ylab="Femur Head Adjusted", main ="Pelvic Breadth vs. Femur Head")

cor(DelPrete$Index_pelvic_breadth, DelPrete$femur_head_adjusted_1.715)
cor(DelPrete$Index_pelvic_breadth, DelPrete$GM)

cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$GM)
cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715)

t.test(DelPreteT$Index_pelvic_breadth ~ DelPreteT$Ethnicity)
```


Since the p-value for the Pelvic Breadth is less than 0.05 so we can conclude that there is a statistical difference between the mean Pelvic Breadth index value of the races of the Terry collection.

# AP Outlet
```{r}

hist(DelPreteT$Index_AP_outlet)
plot(DelPreteTW$Index_AP_outlet, DelPreteTW$GM, xlab = "AP Outlet", ylab="Geometric Mean", main = "AP Outlet vs. Geometric Mean")
plot(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Outlet", ylab="Femur Head Adjusted", main ="AP Outlet vs. Femur Head")

cor(DelPrete$Index_AP_outlet, DelPrete$femur_head_adjusted_1.715)
cor(DelPrete$Index_AP_outlet, DelPrete$GM)


cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715)

View(DelPreteT)
wilcox.test(DelPreteT$Index_AP_outlet ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is less than 0.05 so we can conclude that there is a difference in the ranks of the two races in the Terry collection and therefore there is a difference between the two groups.




