---
title: "R Notebook"
output: html_notebook
---

# 02 24 2020

## Loading the data

```{r}
library(mosaic)
```


```{r}
DelPrete <- read_excel("Delprete.xlsx")

# Changning data into a data frame

DelPrete <- as.data.frame(DelPrete)

View(DelPrete)

library(mosaic)

```

```{r}
DelPrete <- Delprete
DelPreteTW <- subset(DelPreteTW, Collection_number !=1)
DelPreteTW <- subset(DelPreteTW, Collection_number !=2)
DelPreteTW <- subset(DelPreteTW, Collection_number !=3)

# Subsetting the data so its just the Terry Collection

DelPreteT <- subset(DelPrete, Collection_number != 1)
DelPreteT <- subset(DelPreteT, Collection_number != 2)
DelPreteT <- subset(DelPreteT, Collection_number != 3)

```

```{r}
# Subsetting the data so its just the White Terry Collection


DelPreteTW <- subset(DelPrete, Collection_number !=1)
DelPreteTW <- subset(DelPreteTW, Collection_number !=2)
DelPreteTW <- subset(DelPreteTW, Collection_number !=3)
DelPreteTW <- subset(DelPreteTW, Ethnicity !="B")
```

```{r}
sorting.collections.eth <- function(C, E) {
  DelPreteCE <- subset(DelPrete, Collection_number == C)
  DelPreteCE <- subset(DelPreteCE, Ethnicity == E)
  return(DelPreteCE)
}
```

```{r}
sorting.collections.all <- function(C) {
  DelPreteC <- subset(DelPrete, Collection_number == C)
  return(DelPreteC)
}
```

```{r}
DelPreteS <- sorting.collections.all(1)
DelPreteC <- sorting.collections.all(2)
DelPreteHT <- sorting.collections.all(3)
```

## Descriptives

```{r}
favstats(DelPreteTW$Index_AP_inlet)
favstats(DelPreteTW$Index_bispinous)
favstats(DelPreteTW$Index_TD_outlet)
```

# Ap Inlet
```{r}

hist(DelPreteT$Index_AP_inlet, xlab = "AP Inlet", main="Histogram of AP Inlet")
plot(DelPreteTW$Index_AP_inlet, DelPreteTW$GM, xlab = "AP Inlet", ylab="Geometric Mean", main = "AP Inlet vs. Geometric Mean")
plot(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Inlet", ylab="Femur Head Adjusted", main ="AP Inlet vs. Femur Head")

cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715)

t.test(DelPreteT$Index_AP_inlet ~ DelPreteT$Ethnicity)

```


For the t-test, the p-value for the AP Inlet is greater than 0.05 so we cannot conclude that there is a statistical difference between the mean AP Inlet index value of the races of the Terry collection.

# TD Outlet
```{r}

hist(DelPreteT$Index_TD_outlet, xlab="TD Outlet", main="Histogram of TD Outlet")
plot(DelPreteTW$Index_TD_outlet, DelPreteTW$GM, xlab = "TD Outlet", ylab="Geometric Mean", main = "TD Outlet vs. Geometric Mean")
plot(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "TD Outlet", ylab="Femur Head Adjusted", main ="TD Outlet vs. Femur Head")

cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715)

wilcox.test(DelPreteT$Index_TD_outlet ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is greater than 0.05 so we cannot conclude that the ranks for each index value are different and therefore there is no difference between the two groups.


# Bispinous
```{r}

hist(DelPreteT$Index_bispinous, xlab = "Bispinous", main="Histogram of Bispinous")
plot(DelPreteTW$Index_bispinous, DelPreteTW$GM, xlab = "Bispinous", ylab="Geometric Mean", main = "Bispsinous vs. Geometric Mean")
plot(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715, xlab = "Bispinous", ylab="Femur Head Adjusted", main ="Bispinous vs. Femur Head")

cor.test(DelPreteTW$Index_bispinous, DelPreteTW$GM)
cor.test(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715)

wilcox.test(DelPreteT$Index_bispinous ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is greater than 0.05 so we cannot conclude that the ranks for each index value are different and therefore there is no difference between the two groups.

# Pelvic Breadth
```{r}

hist(DelPreteT$Index_pelvic_breadth, xlab = "Pelvic Breadth", main = "Histogram of Pelvic Breadth")
plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$GM, xlab = "Pelvic Breadth", ylab="Geometric Mean", main = "Pelvic Breadth vs. Geometric Mean")
plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715, xlab = "Pelvic Breadth", ylab="Femur Head Adjusted", main ="Pelvic Breadth vs. Femur Head")

cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$GM)
cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715)

t.test(DelPreteT$Index_pelvic_breadth ~ DelPreteT$Ethnicity)
```


Since the p-value for the Pelvic Breadth is less than 0.05 so we can conclude that there is a statistical difference between the mean Pelvic Breadth index value of the races of the Terry collection.

# AP Outlet
```{r}

hist(DelPreteT$Index_AP_outlet, xlab ="AP Outlet", main ="Histogram of AP Outlet")
plot(DelPreteTW$Index_AP_outlet, DelPreteTW$GM, xlab = "AP Outlet", ylab="Geometric Mean", main = "AP Outlet vs. Geometric Mean")
plot(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Outlet", ylab="Femur Head Adjusted", main ="AP Outlet vs. Femur Head")


cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715)

View(DelPreteT)
wilcox.test(DelPreteT$Index_AP_outlet ~ DelPreteT$Ethnicity)
```


For the Mann Whitney U test we see that the p-value is less than 0.05 so we can conclude that there is a difference in the ranks of the two races in the Terry collection and therefore there is a difference between the two groups.



```{r}
install.packages("devtools", dependencies=TRUE)
```
```{r}
library(devtools)
```

```{r}
devtools::install_github("pratikpitreds/nlCor")
```
```{r}
library("nlCor")
```



```{r}
?nlCor
nlCor(DelPreteT$Index_bispinous, DelPreteT$GM, method =c("poly", "expo", "Linear","log"))
plot(DelPreteT$Index_bispinous, DelPreteT$GM)
```

# Correlation Function
```{r}
cor.index <- function(x,y=Delprete$femur_head_adjusted_1.715) {
  nlc <- vector()
  c <- vector()
  Cor.list <- list(nlc, c)
  for (i in 1:length(x)){
    Cor.list$nlc[i] <- c(nlCor(x[[i]],y, method = "log"), nlCor(x[[i]],y, method="poly"), nlCor(x[[i]],y, method = "expo"))
    Cor.list$c[i] <- cor(x[[i]],y)
  }
  return(Cor.list)
}
```

## Cor of All Data
```{r}
x <- list(Delprete$Index_bispinous, Delprete$Index_AP_inlet, Delprete$Index_AP_outlet, Delprete$Index_TD_outlet, Delprete$Index_pelvic_breadth)

cor.index(x, y=Delprete$femur_head_adjusted_1.715)
cor.index(x, y=Delprete$GM)
```

## Cor of T Collection
```{r}
T <- list(DelPreteT$Index_bispinous, DelPreteT$Index_AP_inlet,DelPreteT$Index_AP_outlet, DelPreteT$Index_TD_outlet, DelPreteT$Index_pelvic_breadth)

cor.index(T, y=DelPreteT$femur_head_adjusted_1.715)
cor.index(T, y=DelPreteT$GM)
```

## Cor of S Collection
```{r}
S <- list(DelPreteS$Index_bispinous, DelPreteS$Index_AP_inlet,DelPreteS$Index_AP_outlet, DelPreteS$Index_TD_outlet, DelPreteS$Index_pelvic_breadth)

cor.index(S, y=DelPreteS$femur_head_adjusted_1.715)
cor.index(S, y=DelPreteS$GM)
```

## Cor of HT Collection
```{r}
HT <- list(DelPreteHT$Index_bispinous, DelPreteHT$Index_AP_inlet,DelPreteHT$Index_AP_outlet, DelPreteHT$Index_TD_outlet, DelPreteHT$Index_pelvic_breadth)

cor.index(HT, y=DelPreteHT$femur_head_adjusted_1.715)
cor.index(HT, y=DelPreteHT$GM)
```

## Cor of C Collection
```{r}
C <- list(DelPreteC$Index_bispinous, DelPreteC$Index_AP_inlet,DelPreteC$Index_AP_outlet, DelPreteC$Index_TD_outlet, DelPreteC$Index_pelvic_breadth)

cor.index(C, y=DelPreteC$femur_head_adjusted_1.715)
cor.index(C, y=DelPreteC$GM)
```


