---
title: "DelPrete Scratch Work"
Author: Matthew Lawson
output: html_notebook
---

# 02 25 2020

## Loading the data

```{r}
DelPrete <- read_excel("Delprete.xlsx")

# Changning data into a data frame

DelPrete <- as.data.frame(DelPrete)

DelPrete <- subset(DelPrete, Collection_number != 1)
DelPrete <- subset(DelPrete, Collection_number != 2)
DelPrete <- subset(DelPrete, Collection_number != 3)
DelPrete <- subset(DelPrete, Ethnicity != "B")

View(Delprete)

```



## Histograms 

```{r}
# Histgrams of the different responses

hist(DelPrete$Index_AP_inlet, xlab = "AP Inlet", main = "AP Inlet")
hist(DelPrete$Index_bispinous, xlab = "Bispinous", main = "Bispinous")
hist(DelPrete$Index_TD_outlet, xlab = "TD Outlet", main = "TD Outlet")
```

```{r}
# Histogtrams for predictor variables

hist(DelPrete$GM) # Geometric mean of pelvic size
hist(DelPrete$femur_head_adjusted_1.715) # Size of femur head, used to estimate pelvic size 
hist(DelPrete$three_groups) # Three groups for year of death, Terry collection did not have any pelvic bones from the 1933 group
```





## Favstats for mean, standard deviation, etc

```{r}
library(mosaic) # For favstats


# Responses

favstats(DelPrete$Index_AP_inlet)
favstats(DelPrete$Index_bispinous)
favstats(DelPrete$Index_TD_outlet)
```

```{r}

# Predictors

favstats(DelPrete$GM)
favstats(DelPrete$femur_head_adjusted_1.715)
favstats(DelPrete$Birthdate)
```



## Scatter plots between Responses and predictors

```{r}
plot(DelPrete$Index_AP_inlet, DelPrete$GM, xlab = "AP Inlet", ylab = "Geometric Mean")
plot(DelPrete$Index_bispinous, DelPrete$GM, xlab = "Bispinous", ylab = "Geometric Mean")
plot(DelPrete$Index_TD_outlet, DelPrete$GM, xlab = "TD Outlet", ylab = "Geometric Mean")
```

```{r}
# Scatter plots comparing responses with birth year

plot(DelPrete$Index_AP_inlet, DelPrete$Birthdate)
plot(DelPrete$Index_bispinous, DelPrete$Birthdate)
plot(DelPrete$Index_TD_outlet, DelPrete$Birthdate)
```

```{r}
# Death year

plot(DelPrete$Index_AP_inlet, DelPrete$date_of_death)
plot(DelPrete$Index_bispinous, DelPrete$date_of_death)
plot(DelPrete$Index_TD_outlet, DelPrete$date_of_death)
```

## Correlations

```{r}

# Correlation between responses and Geometric Means

cor(DelPrete$Index_AP_inlet, DelPrete$GM)
cor(DelPrete$Index_bispinous, DelPrete$GM)
cor(DelPrete$Index_TD_outlet, DelPrete$GM)
```

```{r}

#Correlation between responses and femur head adjusted

cor(DelPrete$Index_AP_inlet, DelPrete$femur_head_adjusted_1.715)
cor(DelPrete$Index_bispinous, DelPrete$femur_head_adjusted_1.715)
cor(DelPrete$Index_TD_outlet, DelPrete$femur_head_adjusted_1.715)
```

# 02 26 2020

## Box Plots

```{r}
boxplot(DelPrete$Index_AP_inlet, main = "AP Inlet")
boxplot(DelPrete$Index_bispinous, main = "Bispinous")
boxplot(DelPrete$Index_TD_outlet, main = "TD Outlet")
```

## Skewness

```{r}

library(PerformanceAnalytics)
?skewness
skewness(DelPrete$Index_AP_inlet, method = "sample")
skewness(DelPrete$Index_bispinous, method = "sample")
skewness(DelPrete$Index_TD_outlet, method = "sample")
```

## Kurtosis

```{r}
?kurtosis
kurtosis(DelPrete$Index_AP_inlet)
kurtosis(DelPrete$Index_bispinous)
kurtosis(DelPrete$Index_TD_outlet)
```

# 02 27 2020

## Correlations between death year and pelvic size

```{r}
cor(DelPrete$Index_AP_inlet, DelPrete$date_of_death)
cor(DelPrete$Index_bispinous, DelPrete$date_of_death)
cor(DelPrete$Index_TD_outlet, DelPrete$date_of_death)

```

Correlations are very low. Repsonses are not correlated with the response variables.

```{r}
cor(DelPrete$Index_AP_inlet, DelPrete$Birthdate)
cor(DelPrete$Index_bispinous, DelPrete$Birthdate)
cor(DelPrete$Index_TD_outlet, DelPrete$Birthdate)
```

Birth year and response variables are not correlated at all. Nice.


```{r}
Pelvic.cor <- function(a, b) {
  a <- a
  b <- b
  p.cor <- cor(a, b)
  
  return(p.cor)
}

Pelvic.cor(DelPrete$Index_AP_inlet, DelPrete$Birthdate)
```


