---
title: "DelPrete Final"
Author: Sam Colucci and Matt Lawson
output: html_notebook
---

# Packages

```{r}
library(readxl)
library(mosaic)
library(car)
library(multcomp)
library(lsmeans)
```

# Adjusted Data

```{r}
DelPrete <- read_excel("Delprete.xlsx")

# Changning data into a data frame

DelPrete <- as.data.frame(DelPrete)
```

## Subsetting

```{r}
# Subsetting the data so its just the Terry Collection

DelPreteT <- subset(DelPrete, Collection_number != 1)
DelPreteT <- subset(DelPreteT, Collection_number != 2)
DelPreteT <- subset(DelPreteT, Collection_number != 3)
```

```{r}
# Subsetting the data so its just the White Terry Collection

DelPreteTW <- subset(DelPrete, Collection_number !=1)
DelPreteTW <- subset(DelPreteTW, Collection_number !=2)
DelPreteTW <- subset(DelPreteTW, Collection_number !=3)
```


## AP Inlet

```{r}
hist(DelPreteT$Index_AP_inlet, xlab = "AP Inlet", main="Histogram of AP Inlet")

plot(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Inlet", ylab="Femur Head Adjusted", main ="AP Inlet vs. Femur Head")

cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715)
```

## TD Outlet

```{r}
hist(DelPreteT$Index_TD_outlet, xlab="TD Outlet", main="Histogram of TD Outlet")

plot(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "TD Outlet", ylab="Femur Head Adjusted", main ="TD Outlet vs. Femur Head")

cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715)
```

## Bispinous

```{r}
hist(DelPreteT$Index_bispinous, xlab = "Bispinous", main="Histogram of Bispinous")

plot(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715, xlab = "Bispinous", ylab="Femur Head Adjusted", main ="Bispinous vs. Femur Head")

cor.test(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715)
```

## Pelvic Breadth
```{r}
hist(DelPreteT$Index_pelvic_breadth, xlab = "Pelvic Breadth", main = "Histogram of Pelvic Breadth")

plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715, xlab = "Pelvic Breadth", ylab="Femur Head Adjusted", main ="Pelvic Breadth vs. Femur Head")

cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715)
```

## AP Outlet
```{r}
hist(DelPreteT$Index_AP_outlet, xlab ="AP Outlet", main ="Histogram of AP Outlet")

plot(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Outlet", ylab="Femur Head Adjusted", main ="AP Outlet vs. Femur Head")

cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715)
```


# Raw data

```{r}
DelpreteOriginalData <- read_excel("DelpreteOriginalData.xlsx", 
    sheet = "all data")

# Changning data into a data frame

DelpreteOriginalData <- as.data.frame(DelpreteOriginalData)
```

```{r}
# Changing the collection numbers into Locations

DelPrete.orig <- DelpreteOriginalData

TE <- seq(from =4, to=4, length.out=208)
HT <- seq(from =3, to=3, length.out=240)
C <- seq(from =2, to=2, length.out =155)
S <- seq(from =1, to=1, length.out=68)

Collection <- c(TE,C,S,HT)
DelPrete.orig$Collection <- Collection
DelPrete.orig$Location <- DelPrete.orig$Collection

for (i in 1:length(DelPrete.orig$Collection)) {
  if (DelPrete.orig$Collection[i] == 1){
    DelPrete.orig$Location[i] <- "S"
  }else if (DelPrete.orig$Collection[i] == 2){
    DelPrete.orig$Location[i] <- "C"
  }else if (DelPrete.orig$Collection[i] == 3){
    DelPrete.orig$Location[i] <- "HT"
  }else if (DelPrete.orig$Collection[i] == 4){
    DelPrete.orig$Location[i] <- "T"
  }
}

```

```{r}
# Removing a missing data point
DelPrete.orig <- subset(DelPrete.orig, `femur head diameter` != "NA")
```

```{r}
# Cleaning up the data
# Changing Gender Column so there it's only M or F

for (i in 1:length(DelPrete.orig$Gender)) {
  if (DelPrete.orig$Gender[i] == "male" || DelPrete.orig$Gender[i] == "Male"){
    DelPrete.orig$Gender[i] <- "M"
  } else if (DelPrete.orig$Gender[i] == "female" || DelPrete.orig$Gender[i] == "Female"){
    
    DelPrete.orig$Gender[i] <- "F"
  }
}
```

```{r}
# Cleaning up the data
# Changing Ethnicity Column so there it's only W or B


for (i in 1:length(DelPrete.orig$Ethnicity)) {
  if (DelPrete.orig$Ethnicity[i] == "white" || DelPrete.orig$Ethnicity[i] == "white"){
    DelPrete.orig$Ethnicity[i] <- "W"
  } else if (DelPrete.orig$Ethnicity[i] == "black" || DelPrete.orig$Ethnicity[i] == "Black") {
    DelPrete.orig$Ethnicity[i] <- "B"
  }
  
}

```

```{r}
# Combining the Locations and Ethnicities

DelPrete.orig$LocationR <- DelPrete.orig$Location

for (i in 1:length(DelPrete.orig$Location)) {
  if (DelPrete.orig$Location[i] == "S" && DelPrete.orig$Ethnicity[i] == "W" ){
    DelPrete.orig$LocationR[i] <- "SW"
  }else if (DelPrete.orig$Location[i] == "T" && DelPrete.orig$Ethnicity[i] == "W" ){
    DelPrete.orig$LocationR[i] <- "TW"
  }else if (DelPrete.orig$Location[i] == "T" && DelPrete.orig$Ethnicity[i] == "B" ){
    DelPrete.orig$LocationR[i] <- "TB"
  }else if (DelPrete.orig$Location[i] == "HT" && DelPrete.orig$Ethnicity[i] == "W" ){
    DelPrete.orig$LocationR[i] <- "HTW"
  }else if (DelPrete.orig$Location[i] == "HT" && DelPrete.orig$Ethnicity[i] == "B" ){
    DelPrete.orig$LocationR[i] <- "HTB"
  }else if (DelPrete.orig$Location[i] == "C" && DelPrete.orig$Ethnicity[i] == "W" ){
    DelPrete.orig$LocationR[i] <- "CW"
  }
}

View(DelPrete.orig)
```



```{r}
# Making the predictors into factors for ANCOVA later
DelPrete.orig$Gender <- as.factor(DelPrete.orig$Gender)
DelPrete.orig$LocationR <- as.factor(DelPrete.orig$LocationR)
str(DelPrete.orig)
```





## Renaming

```{r}

# Renaming the variables used to simplify the code

DelPrete.orig$Femur.Head.Adj <- DelPrete.orig$`femur head diameter`^1.715
DelPrete.orig$APInlet <- DelPrete.orig$`Anteroposterior diameter of the inlet`
DelPrete.orig$PelvicBreadth <- DelPrete.orig$`Pelvic breadth`
DelPrete.orig$Bispinous <- DelPrete.orig$`Bi-spinous diameter`
DelPrete.orig$TDOutlet <- DelPrete.orig$`Transverse diameter of the outlet`
DelPrete.orig$APOutlet <- DelPrete.orig$`Anterioposterior space of outlet adj`
```

## Histograms

```{r}
hist(DelPrete.orig$APInlet, xlab = "AP Inlet", main = "AP Inlet", col = "gray")
hist(DelPrete.orig$Bispinous, xlab = "Bispinous", main = "Bispinous", col = "tan")
hist(DelPrete.orig$TDOutlet, xlab = "TD Outlet", main = "TD Outlet", col = "navy")
hist(DelPrete.orig$PelvicBreadth, xlab = "Pelvic Breadth", main = "Pelvic Breadth", col = "purple")
hist(DelPrete.orig$APOutlet, xlab = "AP Outlet", main = "AP Outlet", col = "dark green")
```


## ANCOVA

### AP INLET

```{r}
cor.test(DelPrete.orig$APInlet, DelPrete.orig$Femur.Head.Adj)
```

```{r}
contrasts(DelPrete.orig$LocationR)<-cbind(c(-5,1,1,1,1,1), c(0,-4,1,1,1,1),c(0,0,-3,1,1,1), c(0,0,0,-2,1,1), c(0,0,0,0,-1,1))
model.APInlet<-aov(APInlet ~ Gender*LocationR + Femur.Head.Adj, data = DelPrete.orig)
Anova(model.APInlet, type="III")
```

```{r}
summary.lm(model.APInlet)
```

```{r}
postHocs<-glht(model.APInlet, linfct = mcp(LocationR = "Tukey"))
summary(postHocs)
```

```{r}
confint(postHocs)
```

### AP OUTLET

```{r}
cor.test(DelPrete.orig$APOutlet, DelPrete.orig$Femur.Head.Adj)
```

```{r}
contrasts(DelPrete.orig$LocationR)<-cbind(c(-5,1,1,1,1,1), c(0,-4,1,1,1,1),c(0,0,-3,1,1,1), c(0,0,0,-2,1,1), c(0,0,0,0,-1,1))
model.APOutlet<-aov(APOutlet ~ Gender*LocationR + Femur.Head.Adj, data = DelPrete.orig)
Anova(model.APOutlet, type="III")
```

```{r}
summary.lm(model.APOutlet)
```

```{r}
postHocs<-glht(model.APOutlet, linfct = mcp(LocationR = "Tukey"))
summary(postHocs)
```

```{r}
confint(postHocs)
```

### Bispinous

```{r}
cor.test(DelPrete.orig$Bispinous, DelPrete.orig$Femur.Head.Adj)
```

```{r}
contrasts(DelPrete.orig$LocationR)<-cbind(c(-5,1,1,1,1,1), c(0,-4,1,1,1,1),c(0,0,-3,1,1,1), c(0,0,0,-2,1,1), c(0,0,0,0,-1,1))
model.Bispinous<-aov(Bispinous ~ Gender*LocationR + Femur.Head.Adj, data = DelPrete.orig)
Anova(model.Bispinous, type="III")
```

```{r}
summary.lm(model.Bispinous)
```

```{r}
postHocs<-glht(model.Bispinous, linfct = mcp(LocationR = "Tukey"))
summary(postHocs)
```

```{r}
confint(postHocs)
```

### TD OUTLET

```{r}
cor.test(DelPrete.orig$TDOutlet, DelPrete.orig$Femur.Head.Adj)
```

```{r}
contrasts(DelPrete.orig$LocationR)<-cbind(c(-5,1,1,1,1,1), c(0,-4,1,1,1,1),c(0,0,-3,1,1,1), c(0,0,0,-2,1,1), c(0,0,0,0,-1,1))
model.TDOutlet<-aov(TDOutlet ~ Gender*LocationR + Femur.Head.Adj, data = DelPrete.orig)
Anova(model.TDOutlet, type="III")
```

```{r}
summary.lm(model.TDOutlet)
```

```{r}
postHocs<-glht(model.TDOutlet, linfct = mcp(LocationR = "Tukey"))
summary(postHocs)
```

```{r}
confint(postHocs)
```

### PELVIC BREADTH

```{r}
cor.test(DelPrete.orig$PelvicBreadth, DelPrete.orig$Femur.Head.Adj)
```

```{r}
contrasts(DelPrete.orig$LocationR)<-cbind(c(-5,1,1,1,1,1), c(0,-4,1,1,1,1),c(0,0,-3,1,1,1), c(0,0,0,-2,1,1), c(0,0,0,0,-1,1))
model.PelvicBreadth<-aov(PelvicBreadth ~ Gender*LocationR + Femur.Head.Adj, data = DelPrete.orig)
Anova(model.PelvicBreadth, type="III")
```

```{r}
summary.lm(model.PelvicBreadth)
```

```{r}
postHocs<-glht(model.PelvicBreadth, linfct = mcp(LocationR = "Tukey"))
summary(postHocs)
```

```{r}
confint(postHocs)
```

```{r}
posthocinteractions<-lsmeans(model.PelvicBreadth, pairwise ~ Gender:LocationR, adjust = "Tukey")

posthocinteractions$contrasts

```
# Unadjusted Data

# Organizing Data
```{r}
DelPrete <- All_data_transposed_and_adjusted_spring_2020_stats

# Adding collection number column
T<- seq(from =4, to=4, length.out=208)
HT <- seq(from =3, to=3, length.out=240)
C <- seq(from =2, to=2, length.out =155)
S <- seq(from =1, to=1, length.out=68)

Collection <- c(T,C,S,HT)
DelPrete$Collection <- Collection

# Changing gender in HT collection from "female" & "Female" to just "Female"
Gen <- character(length=60)
Gen[1:60] <- "Female"
DelPrete$Gender[492:551] <- Gen

# Adjusting Femur Head Diameter
fem.adj <- (DelPrete$`femur head diameter`)^1.715
DelPrete$`femur head adjusted` <- fem.adj
```


# Sorting Collections
```{r}
# Sorts samples based on Collection number, Gender and Ethnicity
sorting.collections.eth <- function(C, E, G) {
  DelPreteCE <- subset(DelPrete, Collection == C)
  DelPreteCE <- subset(DelPreteCE, Ethnicity == E)
  DelPreteCE <- subset(DelPreteCE, Gender == G)
  return(DelPreteCE)
}
```

```{r}
# Sorts samples based on Collection number and Gender
sorting.collections.gen <- function(C, G) {
  DelPreteCE <- subset(DelPrete, Collection == C)
  DelPreteCE <- subset(DelPreteCE, Gender == G)
  return(DelPreteCE)
}
```

```{r}
# Sorts samples based on Collection
sorting.collections.all <- function(C) {
  DelPreteC <- subset(DelPrete, Collection == C)
  return(DelPreteC)
}
```

```{r}
# Sorts samples based on Collection number and Ethnicity
sorting.collections.eth.gen <- function(C, E) {
  DelPreteCE <- subset(DelPrete, Collection == C)
  DelPreteCE <- subset(DelPreteCE, Ethnicity == E)
  return(DelPreteCE)
}
```


```{r}
DelPrete.orig.HT <- sorting.collections.all(3)
# All Males and Females in HT

DelPrete.orig.T <- sorting.collections.all(3)
# All Males and Females in T

DelPrete.orig.C <- sorting.collections.all(2)
# Males and Females in C

DelPrete.orig.S <- sorting.collections.all(1)
# Males and Females in S

DelPrete.orig.HTW <- sorting.collections.eth.gen(3,"white")
# White Males and Females in HT
DelPrete.orig.HTB <- sorting.collections.eth.gen(3, "Black")
# Black Males and Females in HT

DelPrete.orig.TW <- sorting.collections.eth.gen(4,"W")
# White Males and Females in T
DelPrete.orig.TB <- sorting.collections.eth.gen(4,"B")
# Black Males and Females in T

DelPrete.orig.SF <- sorting.collections.eth(1, "W", "F") 
# White Females in S Collection
DelPrete.orig.SM <- sorting.collections.eth(1, "W", "M")
# White Males in S Collection

DelPrete.orig.CF <- sorting.collections.eth(2, "W", "F")
# White Females in C Collection
DelPrete.orig.CM <- sorting.collections.eth(2, "W", "M")
# White Males in C Collection

DelPrete.orig.TF <- sorting.collections.gen(4, "F")
# Black and White Females in T Collection
DelPrete.orig.TM <- sorting.collections.gen(4, "M")
# Black and White Males in T Collection

DelPrete.orig.HTF <- sorting.collections.gen(3, "Female")
# Black and White Females in HT Collection
DelPrete.orig.HTM <- sorting.collections.gen(3, "Male")
# Black and White Males in HT Collection

```
*** S collection only has 36 samples (smaller than others)


```{r}
DelPrete.orig.HTBF <- sorting.collections.eth(3,"Black","Female")
# Black Females in HT Collection
DelPrete.orig.HTBM <- sorting.collections.eth(3,"Black","Male")
# Black Males in HT Collection

DelPrete.orig.HTWF <- sorting.collections.eth(3,"white","Female")
# White Females in HT Collection
DelPrete.orig.HTWM <- sorting.collections.eth(3,"white","Male")
# White Males in HT Collection

DelPrete.orig.TBF <- sorting.collections.eth(4, "B","F")
# Black Females in T Collection
DelPrete.orig.TBM <- sorting.collections.eth(4, "B","M")
# Black Males in T Collection

DelPrete.orig.TWF <- sorting.collections.eth(4,"W","F")
# White Females in T Collection
DelPrete.orig.TWM <- sorting.collections.eth(4,"W","M")
# White Males in T Collection
```


# Variability
```{r}
# Function to compute coefficient of variation
var.test <- function(x) {
  var <- sd(x)*100/mean(x)
  return(var)
}
```

```{r}
# Install asymptotic tests
install.packages("cvequality")
library(cvequality)
```

## AP Inlet Coefficients of Variation
```{r}
# HTW - Female
var.ap.f.htw <- var.test(DelPrete.orig.HTWF$`Anteroposterior diameter of the inlet`)

# HTW - Male 
var.ap.m.htw <- var.test(DelPrete.orig.HTWM$`Anteroposterior diameter of the inlet`)

# HTB - Female
var.ap.f.htb <- var.test(DelPrete.orig.HTBF$`Anteroposterior diameter of the inlet`)

# HTB - Male
var.ap.m.htb <- var.test(DelPrete.orig.HTBM$`Anteroposterior diameter of the inlet`)

# TW - Female
var.ap.f.tw <- var.test(DelPrete.orig.TWF$`Anteroposterior diameter of the inlet`)

# TW -  Male
var.ap.m.tw <- var.test(DelPrete.orig.TWM$`Anteroposterior diameter of the inlet`)

# TB - Female
var.ap.f.tb <- var.test(DelPrete.orig.TBF$`Anteroposterior diameter of the inlet`)

# TB - Male
var.ap.m.tb <- var.test(DelPrete.orig.TBM$`Anteroposterior diameter of the inlet`)

# C - Female
var.ap.f.c <- var.test(DelPrete.orig.CF$`Anteroposterior diameter of the inlet`)

# C - Male
var.ap.m.c <- var.test(DelPrete.orig.CM$`Anteroposterior diameter of the inlet`)

# S - Female
var.ap.f.s <- var.test(DelPrete.orig.SF$`Anteroposterior diameter of the inlet`)

# S - Male
var.ap.m.s <- var.test(DelPrete.orig.SM$`Anteroposterior diameter of the inlet`)
```

## Pelvic Breadth Coefficients of Variation

```{r}
# HTW - Female
var.pb.f.htw <- var.test(DelPrete.orig.HTWF$`Pelvic breadth`)

# HTW - Male 
var.pb.m.htw <- var.test(DelPrete.orig.HTWM$`Pelvic breadth`)

# HTB - Female
var.pb.f.htb <- var.test(DelPrete.orig.HTBF$`Pelvic breadth`)

# HTB - Male
var.pb.m.htb <- var.test(DelPrete.orig.HTBM$`Pelvic breadth`)

# TW - Female
var.pb.f.tw <- var.test(DelPrete.orig.TWF$`Pelvic breadth`)

# TW - Male
var.pb.m.tw <- var.test(DelPrete.orig.TWM$`Pelvic breadth`)

# TB - Female
var.pb.f.tb <- var.test(DelPrete.orig.TBF$`Pelvic breadth`)

# TB - Male
var.pb.m.tb <- var.test(DelPrete.orig.TBM$`Pelvic breadth`)

# C - Female
var.pb.f.c <- var.test(DelPrete.orig.CF$`Pelvic breadth`)

# C - Male
var.pb.m.c <- var.test(DelPrete.orig.CM$`Pelvic breadth`)

# S - Female
var.pb.f.s <- var.test(DelPrete.orig.SF$`Pelvic breadth`)

# S - Male
var.pb.m.s <- var.test(DelPrete.orig.SM$`Pelvic breadth`)
```

## Bispinous Coefficients of Variation

```{r}
# HTW - Female
var.b.f.htw <- var.test(DelPrete.orig.HTWF$`Bi-spinous diameter`)

# HTW - Male
var.b.m.htw <- var.test(DelPrete.orig.HTWM$`Bi-spinous diameter`)

# HTB - Female
var.b.f.htb <- var.test(DelPrete.orig.HTBF$`Bi-spinous diameter`)

# HTB - Male
var.b.m.htb <- var.test(DelPrete.orig.HTBM$`Bi-spinous diameter`)

# TW - Female
var.b.f.tw <- var.test(DelPrete.orig.TWF$`Bi-spinous diameter`)

# TW - Male
var.b.m.tw <- var.test(DelPrete.orig.TWM$`Bi-spinous diameter`)

# TB - Female
var.b.f.tb <- var.test(DelPrete.orig.TBF$`Bi-spinous diameter`)

# TB - Male
var.b.m.tb <- var.test(DelPrete.orig.TBM$`Bi-spinous diameter`)

# C - Female
var.b.f.c <- var.test(DelPrete.orig.CF$`Bi-spinous diameter`)

# C - Male
c <- var.test(DelPrete.orig.CM$`Bi-spinous diameter`)

# S - Female
var.b.f.s <- var.test(DelPrete.orig.SF$`Bi-spinous diameter`)

# S - Male
var.b.m.s <- var.test(DelPrete.orig.SM$`Bi-spinous diameter`)
```

## AP Outlet Coefficients of Variation

```{r}
# HTW - Female
var.apo.f.htw <- var.test(DelPrete.orig.HTWF$`Anterioposterior space of outlet adj`)

# HTW - Male
var.apo.m.htw <- var.test(DelPrete.orig.HTWM$`Anterioposterior space of outlet adj`)

# HTB - Female
var.apo.f.htb <- var.test(DelPrete.orig.HTBF$`Anterioposterior space of outlet adj`)

# HTB - Male
var.apo.m.htb <- var.test(DelPrete.orig.HTBM$`Anterioposterior space of outlet adj`)

# TW - Female
var.apo.f.tw <- var.test(DelPrete.orig.TWF$`Anterioposterior space of outlet adj`)

# TW - Male
var.apo.m.tw <- var.test(DelPrete.orig.TWM$`Anterioposterior space of outlet adj`)

# TB - Female
var.apo.f.tb <- var.test(DelPrete.orig.TBF$`Anterioposterior space of outlet adj`)

# TB - Male
var.apo.m.tb <- var.test(DelPrete.orig.TBM$`Anterioposterior space of outlet adj`)

# C - Female
var.apo.f.c <- var.test(DelPrete.orig.CF$`Anterioposterior space of outlet adj`)

# C - Male
var.apo.m.c <- var.test(DelPrete.orig.CM$`Anterioposterior space of outlet adj`)

# S - Female
var.apo.f.s <- var.test(DelPrete.orig.SF$`Anterioposterior space of outlet adj`)

# S - Male
var.apo.m.s <- var.test(DelPrete.orig.SM$`Anterioposterior space of outlet adj`)
```

## TD Outlet Coefficients of Variation

```{r}
# HTW - Female
var.td.f.htw <- var.test(DelPrete.orig.HTWF$`Transverse diameter of the outlet`)

# HTW - Male
var.td.m.htw <- var.test(DelPrete.orig.HTWM$`Transverse diameter of the outlet`)

# HTB - Female
var.td.f.htb <- var.test(DelPrete.orig.HTBF$`Transverse diameter of the outlet`)

# HTB - Male
var.td.m.htb <- var.test(DelPrete.orig.HTBM$`Transverse diameter of the outlet`)

# TW - Female
var.td.f.tw <- var.test(DelPrete.orig.TWF$`Transverse diameter of the outlet`)

# TW - Male
var.td.m.tw <- var.test(DelPrete.orig.TWM$`Transverse diameter of the outlet`)

# TB - Female
var.td.f.tb <- var.test(DelPrete.orig.TBF$`Transverse diameter of the outlet`)

# TB - Male
var.td.m.tb <- var.test(DelPrete.orig.TBM$`Transverse diameter of the outlet`)

# C - Female
var.td.f.c <- var.test(DelPrete.orig.CF$`Transverse diameter of the outlet`)

# C - Male
var.td.m.c <- var.test(DelPrete.orig.CM$`Transverse diameter of the outlet`)

# S - Female
var.td.f.s <- var.test(DelPrete.orig.SF$`Transverse diameter of the outlet`)

# S - Male
var.td.m.s <- var.test(DelPrete.orig.SM$`Transverse diameter of the outlet`)
```

## Variability Matrices by Index Value
```{r}
# AP Inlet
var.mat.ap <- matrix(c(var.ap.f.htw,var.ap.f.htb,var.ap.f.tw, var.ap.f.tb, var.ap.f.c, var.ap.f.s, var.ap.m.htw, var.ap.m.htb, var.ap.m.tw, var.ap.m.tb, var.ap.m.c, var.ap.m.s),nrow=6, ncol=2)

# AP Outlet
var.mat.apo <- matrix(c(var.apo.f.htw,var.apo.f.htb,var.apo.f.tw, var.apo.f.tb, var.apo.f.c, var.apo.f.s, var.apo.m.htw, var.apo.m.htb, var.apo.m.tw, var.apo.m.tb, var.apo.m.c, var.apo.m.s),nrow=6, ncol=2)

# TD Outlet
var.mat.td <- matrix(c(var.td.f.htw,var.td.f.htb,var.td.f.tw, var.td.f.tb, var.td.f.c, var.td.f.s, var.td.m.htw, var.td.m.htb, var.td.m.tw, var.td.m.tb, var.td.m.c, var.td.m.s),nrow=6, ncol=2)

# Pelvic Breadth
var.mat.pb <- matrix(c(var.pb.f.htw,var.pb.f.htb,var.pb.f.tw, var.pb.f.tb, var.pb.f.c, var.pb.f.s, var.pb.m.htw, var.pb.m.htb, var.pb.m.tw, var.pb.m.tb, var.pb.m.c, var.pb.m.s),nrow=6, ncol=2)

# Bispinous
var.mat.b <- matrix(c(var.b.f.htw,var.b.f.htb,var.b.f.tw, var.b.f.tb, var.b.f.c, var.b.f.s, var.b.m.htw, var.b.m.htb, var.b.m.tw, var.b.m.tb, var.b.m.c, var.b.m.s),nrow=6, ncol=2)

```

## Var Matrices by Collection
```{r}
# HTW
htw.var.mat <- matrix(c(var.ap.f.htw, var.apo.f.htw, var.td.f.htw, var.pb.f.htw, var.b.f.htw, var.ap.m.htw, var.apo.m.htw, var.td.m.htw, var.pb.m.htw, var.b.m.htw), nrow=5)

# HTB
htb.var.mat <- matrix(c(var.ap.f.htb, var.apo.f.htb, var.td.f.htb, var.pb.f.htb, var.b.f.htb, var.ap.m.htb, var.apo.m.htb, var.td.m.htb, var.pb.m.htb, var.b.m.htb), nrow=5)

# TW
tw.var.mat <- matrix(c(var.ap.f.tw, var.apo.f.tw, var.td.f.tw, var.pb.f.tw, var.b.f.tw, var.ap.m.tw, var.apo.m.tw, var.td.m.tw, var.pb.m.tw, var.b.m.tw), nrow=5)

# TB
tb.var.mat <- matrix(c(var.ap.f.tb, var.apo.f.tb, var.td.f.tb, var.pb.f.tb, var.b.f.tb, var.ap.m.tb, var.apo.m.tb, var.td.m.tb, var.pb.m.tb, var.b.m.tb), nrow=5)

# C
c.var.mat <- matrix(c(var.ap.f.c, var.apo.f.c, var.td.f.c, var.pb.f.c, var.b.f.c, var.ap.m.c, var.apo.m.c, var.td.m.c, var.pb.m.c, var.b.m.c), nrow=5)

# S
s.var.mat <- matrix(c(var.ap.f.s, var.apo.f.s, var.td.f.s, var.pb.f.s, var.b.f.s, var.ap.m.s, var.apo.m.s, var.td.m.s, var.pb.m.s, var.b.m.s), nrow=5)
```


## Table and List of Variability
```{r}
# Matrix of all Coefficients of Variability
var.mat <- matrix(c(var.mat.ap, var.mat.apo, var.mat.td, var.mat.pb, var.mat.b), nrow=6)

# Vector of all female Cv
female.vals <- c(var.mat[1:6,1], var.mat[1:6,3], var.mat[1:6, 5], var.mat[1:6,7], var.mat[1:6,9])

# Vector of all male Cv
male.vals <- c(var.mat[1:6,2], var.mat[1:6,4], var.mat[1:6,6], var.mat[1:6,8], var.mat[1:6,10])

var.mat.final <- matrix(c(female.vals, male.vals), nrow=30)

# Tables of Matrices
ap.table <- as.table(var.mat.final[1:6,])
colnames(ap.table) = c("Female", "Male")
rownames(ap.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

apo.table <- as.table(var.mat.final[7:12,])
colnames(apo.table) = c("Female", "Male")
rownames(apo.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

td.table <- as.table(var.mat.final[13:18,])
colnames(td.table) = c("Female", "Male")
rownames(td.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

pb.table <- as.table(var.mat.final[19:24,])
colnames(pb.table) = c("Female", "Male")
rownames(pb.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

b.table <- as.table(var.mat.final[25:30,])
colnames(b.table) = c("Female", "Male")
rownames(b.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

# List of Cvs
var.list <- list("AP Inlet" = ap.table,"AP Outlet" = apo.table, "TD Outlet" = td.table, "Pelvic Breadth" = pb.table, "Bispinous" = b.table)
```

## Significance of Variability

### AP Inlet
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.ap <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.HTB.ap <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.TW.ap <- with(DelPrete.orig.TW, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.TB.ap <- with(DelPrete.orig.TB, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.C.ap <- with(DelPrete.orig.C, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.S.ap <- with(DelPrete.orig.S, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

# Vector of Sig. Levels
var.sig.levels.ap <- c(sig.diff.HTW.ap$p_value, sig.diff.HTB.ap$p_value, sig.diff.TW.ap$p_value, sig.diff.TB.ap$p_value, sig.diff.C.ap$p_value, sig.diff.S.ap$p_value)

# AP Table w/ Sig Levels
ap.table.new <- as.table(matrix(c(var.mat.ap, var.sig.levels.ap), nrow=6))
colnames(ap.table.new) = c("Female", "Male", "Pvalue")
row.names(ap.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# AP Dataframe w/ Sig Levels
ap.mat <- matrix(c(var.mat.ap, var.sig.levels.ap), nrow=6)
ap.df <- as.data.frame(ap.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(ap.df)= c("Female", "Male", "P-value")
ap.df
```

### AP Outlet
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.apo <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.HTB.apo <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.TW.apo <- with(DelPrete.orig.TW, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.TB.apo <- with(DelPrete.orig.TB, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.C.apo <- with(DelPrete.orig.C, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.S.apo <- with(DelPrete.orig.S, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

# Vector of Sig Levels
var.sig.levels.apo <- c(sig.diff.HTW.apo$p_value, sig.diff.HTB.apo$p_value, sig.diff.TW.apo$p_value, sig.diff.TB.apo$p_value, sig.diff.C.apo$p_value, sig.diff.S.apo$p_value)

# APO Table w/ Sig Levels
apo.table.new <- as.table(matrix(c(var.mat.apo, var.sig.levels.apo), nrow=6))
colnames(apo.table.new) = c("Female", "Male", "Pvalue")
row.names(apo.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# APO Dataframe w/ Sig Levels
apo.mat <- matrix(c(var.mat.apo, var.sig.levels.apo), nrow=6)
apo.df <- as.data.frame(apo.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(apo.df)= c("Female", "Male", "P-value")
apo.df
```

### TD Outlet
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.td <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.HTB.td <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.TW.td <- with(DelPrete.orig.TW, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.TB.td <- with(DelPrete.orig.TB, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.C.td <- with(DelPrete.orig.C, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.S.td <- with(DelPrete.orig.S, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

# Vector of Sig Levels
var.sig.levels.td <- c(sig.diff.HTW.td$p_value, sig.diff.HTB.td$p_value, sig.diff.TW.td$p_value, sig.diff.TB.td$p_value, sig.diff.C.td$p_value, sig.diff.S.td$p_value)

# TD Table w/ Sig Levels
td.table.new <- as.table(matrix(c(var.mat.td, var.sig.levels.td), nrow=6))
colnames(td.table.new) = c("Female", "Male", "Pvalue")
row.names(td.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# TD Dataframe w/ Sig Levels
td.mat <- matrix(c(var.mat.td, var.sig.levels.td), nrow=6)
td.df <- as.data.frame(td.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(td.df)= c("Female", "Male", "P-value")
td.df
```

### Pelvic Breadth
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.pb <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.HTB.pb <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.TW.pb <- with(DelPrete.orig.TW, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.TB.pb <- with(DelPrete.orig.TB, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.C.pb <- with(DelPrete.orig.C, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.S.pb <- with(DelPrete.orig.S, 
     asymptotic_test(`Pelvic breadth`,Gender))

# Vector of Sig Levels
var.sig.levels.pb <- c(sig.diff.HTW.pb$p_value, sig.diff.HTB.pb$p_value, sig.diff.TW.pb$p_value, sig.diff.TB.pb$p_value, sig.diff.C.pb$p_value, sig.diff.S.pb$p_value)

# PB Table w/ Sig Levels
pb.table.new <- as.table(matrix(c(var.mat.pb, var.sig.levels.pb), nrow=6))
colnames(pb.table.new) = c("Female", "Male", "Pvalue")
row.names(pb.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# PB Dataframe w/ Sig Levels
pb.mat <- matrix(c(var.mat.pb, var.sig.levels.pb), nrow=6)
pb.df <- as.data.frame(pb.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(pb.df)= c("Female", "Male", "P-value")
pb.df
```

### Bispinous
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.b <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.HTB.b <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.TW.b <- with(DelPrete.orig.TW, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.TB.b <- with(DelPrete.orig.TB, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.C.b <- with(DelPrete.orig.C, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.S.b <- with(DelPrete.orig.S, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

# Vector of Sig Levels
var.sig.levels.b <- c(sig.diff.HTW.b$p_value, sig.diff.HTB.b$p_value, sig.diff.TW.b$p_value, sig.diff.TB.b$p_value, sig.diff.C.b$p_value, sig.diff.S.b$p_value)

# B Table w/ Sig Levels
b.table.new <- as.table(matrix(c(var.mat.b, var.sig.levels.b), nrow=6))
colnames(b.table.new) = c("Female", "Male", "Pvalue")
row.names(b.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# B Dataframe w/ Sig Levels
b.mat <- matrix(c(var.mat.b, var.sig.levels.b), nrow=6)
b.df <- as.data.frame(b.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(b.df)= c("Female", "Male", "P-value")
b.df
```

## New Variations List
```{r}
# List of Cvs w/ Sig Levels
var.list.new <- list("AP Inlet" = ap.table.new,"AP Outlet" = apo.table.new, "TD Outlet" = td.table.new, "Pelvic Breadth" = pb.table.new, "Bispinous" = b.table.new)

# List of Cv Dataframes w/ Sig Levels
final.var.list <- list("AP Inlet"=ap.df, "AP Outlet"=apo.df, "TD Outlet"=td.df, "Pelvic Breadth"=pb.df, "Bispinous"=b.df)
final.var.list
```
Using 0.05 conf. level: C - AP Outlet, C - Pelvic Breadth have significant difference
Using 0.1 conf. level: HTB, TW, C, S - AP Outlet, HTB, C - Pelvic Breadth 
Very small amount of collections by index value have a significant difference between males and females

