---
title: "R Notebook"
output: html_notebook
---

# 02 24 2020

# Loading the data

```{r}
library(mosaic)

```


```{r}
DelPrete <- read_excel("Delprete.xlsx")

# Changning data into a data frame

DelPrete <- as.data.frame(DelPrete)

View(DelPrete)

library(mosaic)

```

```{r}

DelPreteTW <- subset(DelPrete, Collection_number !=1)
DelPreteTW <- subset(DelPreteTW, Collection_number !=2)
DelPreteTW <- subset(DelPreteTW, Collection_number !=3)

# Subsetting the data so its just the Terry Collection

DelPreteT <- subset(DelPrete, Collection_number != 1)
DelPreteT <- subset(DelPreteT, Collection_number != 2)
DelPreteT <- subset(DelPreteT, Collection_number != 3)

```

```{r}
# Subsetting the data so its just the White Terry Collection


DelPreteTW <- subset(DelPrete, Collection_number !=1)
DelPreteTW <- subset(DelPreteTW, Collection_number !=2)
DelPreteTW <- subset(DelPreteTW, Collection_number !=3)
DelPreteTW <- subset(DelPreteTW, Ethnicity !="B")
```

```{r}
sorting.collections.eth <- function(C, E) {
  DelPreteCE <- subset(DelPrete, Collection_number == C)
  DelPreteCE <- subset(DelPreteCE, Ethnicity == E)
  return(DelPreteCE)
}
```

```{r}
sorting.collections.all <- function(C) {
  DelPreteC <- subset(DelPrete, Collection_number == C)
  return(DelPreteC)
}
```

```{r}
DelPreteS <- sorting.collections.all(1)
DelPreteC <- sorting.collections.all(2)
DelPreteHT <- sorting.collections.all(3)
```

# Descriptives

```{r}
favstats(DelPreteTW$Index_AP_inlet)
favstats(DelPreteTW$Index_bispinous)
favstats(DelPreteTW$Index_TD_outlet)
```

## Ap Inlet
```{r}

hist(DelPreteT$Index_AP_inlet, xlab = "AP Inlet", main="Histogram of AP Inlet")
plot(DelPreteTW$Index_AP_inlet, DelPreteTW$GM, xlab = "AP Inlet", ylab="Geometric Mean", main = "AP Inlet vs. Geometric Mean")
plot(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Inlet", ylab="Femur Head Adjusted", main ="AP Inlet vs. Femur Head")

cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_AP_inlet, DelPreteTW$femur_head_adjusted_1.715)

t.test(DelPreteT$Index_AP_inlet ~ DelPreteT$Ethnicity)

```

For the t-test, the p-value for the AP Inlet is greater than 0.05 so we cannot conclude that there is a statistical difference between the mean AP Inlet index value of the races of the Terry collection. This conclusion is also supposrted by the confidence interval.

## TD Outlet
```{r}

hist(DelPreteT$Index_TD_outlet, xlab="TD Outlet", main="Histogram of TD Outlet")
plot(DelPreteTW$Index_TD_outlet, DelPreteTW$GM, xlab = "TD Outlet", ylab="Geometric Mean", main = "TD Outlet vs. Geometric Mean")
plot(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "TD Outlet", ylab="Femur Head Adjusted", main ="TD Outlet vs. Femur Head")

cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_TD_outlet, DelPreteTW$femur_head_adjusted_1.715)

wilcox.test(DelPreteT$Index_TD_outlet ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is greater than 0.05 so we cannot conclude that the ranks for each index value are different and therefore there is no difference between the two groups. This conclusion is also supported by the confidence interval.


## Bispinous
```{r}

hist(DelPreteT$Index_bispinous, xlab = "Bispinous", main="Histogram of Bispinous")
plot(DelPreteTW$Index_bispinous, DelPreteTW$GM, xlab = "Bispinous", ylab="Geometric Mean", main = "Bispsinous vs. Geometric Mean")
plot(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715, xlab = "Bispinous", ylab="Femur Head Adjusted", main ="Bispinous vs. Femur Head")

cor.test(DelPreteTW$Index_bispinous, DelPreteTW$GM)
cor.test(DelPreteTW$Index_bispinous, DelPreteTW$femur_head_adjusted_1.715)

wilcox.test(DelPreteT$Index_bispinous ~ DelPreteT$Ethnicity)
```

For the Wilcoxson-Signed Rank Test we see that the p-value is greater than 0.05 so we cannot conclude that the ranks for each index value are different and therefore there is no difference between the two groups. This is also supported by the confidence interval.

## Pelvic Breadth
```{r}

hist(DelPreteT$Index_pelvic_breadth, xlab = "Pelvic Breadth", main = "Histogram of Pelvic Breadth")
plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$GM, xlab = "Pelvic Breadth", ylab="Geometric Mean", main = "Pelvic Breadth vs. Geometric Mean")
plot(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715, xlab = "Pelvic Breadth", ylab="Femur Head Adjusted", main ="Pelvic Breadth vs. Femur Head")

cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$GM)
cor.test(DelPreteTW$Index_pelvic_breadth, DelPreteTW$femur_head_adjusted_1.715)

t.test(DelPreteT$Index_pelvic_breadth ~ DelPreteT$Ethnicity)
```


Since the p-value for the Pelvic Breadth is less than 0.05 so we can conclude that there is a statistical difference between the mean Pelvic Breadth index value of the races of the Terry collection. This is supported by the confidence interval not containing 0.

## AP Outlet
```{r}

hist(DelPreteT$Index_AP_outlet, xlab ="AP Outlet", main ="Histogram of AP Outlet")
plot(DelPreteTW$Index_AP_outlet, DelPreteTW$GM, xlab = "AP Outlet", ylab="Geometric Mean", main = "AP Outlet vs. Geometric Mean")
plot(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715, xlab = "AP Outlet", ylab="Femur Head Adjusted", main ="AP Outlet vs. Femur Head")


cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$GM)
cor.test(DelPreteTW$Index_AP_outlet, DelPreteTW$femur_head_adjusted_1.715)

View(DelPreteT)
wilcox.test(DelPreteT$Index_AP_outlet ~ DelPreteT$Ethnicity)
```


For the Mann Whitney U test we see that the p-value is less than 0.05 so we can conclude that there is a difference in the ranks of the two races in the Terry collection and therefore there is a difference between the two groups.



```{r}
install.packages("devtools", dependencies=TRUE)
```
```{r}
library(devtools)
```

```{r}
devtools::install_github("pratikpitreds/nlCor")
```
```{r}
library("nlCor")
```



```{r}
?nlCor
nlCor(DelPreteT$Index_bispinous, DelPreteT$GM, method =c("poly", "expo", "Linear","log"))
plot(DelPreteT$Index_bispinous, DelPreteT$GM)
```

# Correlation Function
```{r}
cor.index <- function(x,y) {
  nlc <- vector()
  c <- vector()
  Cor.list <- list(nlc, c)
  for (i in 1:length(x)){
    Cor.list$nlc[i] <- c(nlCor(x[[i]],y, method = "log", n=3), nlCor(x[[i]],y, method="poly", n=3), nlCor(x[[i]],y, method = "expo", n=3))
    Cor.list$c[i] <- cor(x[[i]],y)
  }
  return(Cor.list)
}
```

## Cor of All Data
```{r}
x <- list(Delprete$Index_bispinous, Delprete$Index_AP_inlet, Delprete$Index_AP_outlet, Delprete$Index_TD_outlet, Delprete$Index_pelvic_breadth)

cor.index(x, y=Delprete$femur_head_adjusted_1.715)
cor.index(x, y=Delprete$GM)

```

## Cor of T Collection
```{r}
T <- list(DelPreteT$Index_bispinous, DelPreteT$Index_AP_inlet,DelPreteT$Index_AP_outlet, DelPreteT$Index_TD_outlet, DelPreteT$Index_pelvic_breadth)

cor.index(T, y=DelPreteT$femur_head_adjusted_1.715)
cor.index(T, y=DelPreteT$GM)
```

## Cor of S Collection
```{r}
S <- list(DelPreteS$Index_bispinous, DelPreteS$Index_AP_inlet,DelPreteS$Index_AP_outlet, DelPreteS$Index_TD_outlet, DelPreteS$Index_pelvic_breadth)

cor.index(S, y=DelPreteS$femur_head_adjusted_1.715)
cor.index(S, y=DelPreteS$GM)
```

## Cor of HT Collection
```{r}
HT <- list(DelPreteHT$Index_bispinous, DelPreteHT$Index_AP_inlet,DelPreteHT$Index_AP_outlet, DelPreteHT$Index_TD_outlet, DelPreteHT$Index_pelvic_breadth)

cor.index(HT, y=DelPreteHT$femur_head_adjusted_1.715)
cor.index(HT, y=DelPreteHT$GM)
```

## Cor of C Collection
```{r}
C <- list(DelPreteC$Index_bispinous, DelPreteC$Index_AP_inlet,DelPreteC$Index_AP_outlet, DelPreteC$Index_TD_outlet, DelPreteC$Index_pelvic_breadth)

cor.index(C, y=DelPreteC$femur_head_adjusted_1.715)
cor.index(C, y=DelPreteC$GM)
```

# Unadjusted Data

# Organizing Data
```{r}
DelPrete <- All_data_transposed_and_adjusted_spring_2020_stats

# Adding collection number column
T<- seq(from =4, to=4, length.out=208)
HT <- seq(from =3, to=3, length.out=240)
C <- seq(from =2, to=2, length.out =155)
S <- seq(from =1, to=1, length.out=68)

Collection <- c(T,C,S,HT)
DelPrete$Collection <- Collection

# Changing gender in HT collection from "female" & "Female" to just "Female"
Gen <- character(length=60)
Gen[1:60] <- "Female"
DelPrete$Gender[492:551] <- Gen

# Adjusting Femur Head Diameter
fem.adj <- (DelPrete$`femur head diameter`)^1.715
DelPrete$`femur head adjusted` <- fem.adj
```


# Sorting Collections
```{r}
# Sorts samples based on Collection number, Gender and Ethnicity
sorting.collections.eth <- function(C, E, G) {
  DelPreteCE <- subset(DelPrete, Collection == C)
  DelPreteCE <- subset(DelPreteCE, Ethnicity == E)
  DelPreteCE <- subset(DelPreteCE, Gender == G)
  return(DelPreteCE)
}
```

```{r}
# Sorts samples based on Collection number and Gender
sorting.collections.gen <- function(C, G) {
  DelPreteCE <- subset(DelPrete, Collection == C)
  DelPreteCE <- subset(DelPreteCE, Gender == G)
  return(DelPreteCE)
}
```

```{r}
# Sorts samples based on Collection
sorting.collections.all <- function(C) {
  DelPreteC <- subset(DelPrete, Collection == C)
  return(DelPreteC)
}
```

```{r}
# Sorts samples based on Collection number and Ethnicity
sorting.collections.eth.gen <- function(C, E) {
  DelPreteCE <- subset(DelPrete, Collection == C)
  DelPreteCE <- subset(DelPreteCE, Ethnicity == E)
  return(DelPreteCE)
}
```


```{r}
DelPrete.orig.HT <- sorting.collections.all(3)
# All Males and Females in HT

DelPrete.orig.T <- sorting.collections.all(3)
# All Males and Females in T

DelPrete.orig.C <- sorting.collections.all(2)
# Males and Females in C

DelPrete.orig.S <- sorting.collections.all(1)
# Males and Females in S

DelPrete.orig.HTW <- sorting.collections.eth.gen(3,"white")
# White Males and Females in HT
DelPrete.orig.HTB <- sorting.collections.eth.gen(3, "Black")
# Black Males and Females in HT

DelPrete.orig.TW <- sorting.collections.eth.gen(4,"W")
# White Males and Females in T
DelPrete.orig.TB <- sorting.collections.eth.gen(4,"B")
# Black Males and Females in T

DelPrete.orig.SF <- sorting.collections.eth(1, "W", "F") 
# White Females in S Collection
DelPrete.orig.SM <- sorting.collections.eth(1, "W", "M")
# White Males in S Collection

DelPrete.orig.CF <- sorting.collections.eth(2, "W", "F")
# White Females in C Collection
DelPrete.orig.CM <- sorting.collections.eth(2, "W", "M")
# White Males in C Collection

DelPrete.orig.TF <- sorting.collections.gen(4, "F")
# Black and White Females in T Collection
DelPrete.orig.TM <- sorting.collections.gen(4, "M")
# Black and White Males in T Collection

DelPrete.orig.HTF <- sorting.collections.gen(3, "Female")
# Black and White Females in HT Collection
DelPrete.orig.HTM <- sorting.collections.gen(3, "Male")
# Black and White Males in HT Collection

```
*** S collection only has 36 samples (smaller than others)


```{r}
DelPrete.orig.HTBF <- sorting.collections.eth(3,"Black","Female")
# Black Females in HT Collection
DelPrete.orig.HTBM <- sorting.collections.eth(3,"Black","Male")
# Black Males in HT Collection

DelPrete.orig.HTWF <- sorting.collections.eth(3,"white","Female")
# White Females in HT Collection
DelPrete.orig.HTWM <- sorting.collections.eth(3,"white","Male")
# White Males in HT Collection

DelPrete.orig.TBF <- sorting.collections.eth(4, "B","F")
# Black Females in T Collection
DelPrete.orig.TBM <- sorting.collections.eth(4, "B","M")
# Black Males in T Collection

DelPrete.orig.TWF <- sorting.collections.eth(4,"W","F")
# White Females in T Collection
DelPrete.orig.TWM <- sorting.collections.eth(4,"W","M")
# White Males in T Collection
```


# Variability
```{r}
# Function to compute coefficient of variation
var.test <- function(x) {
  var <- sd(x)*100/mean(x)
  return(var)
}
```

```{r}
# Install asymptotic tests
install.packages("cvequality")
library(cvequality)
```

## AP Inlet Coefficients of Variation
```{r}
# HTW - Female
var.ap.f.htw <- var.test(DelPrete.orig.HTWF$`Anteroposterior diameter of the inlet`)

# HTW - Male 
var.ap.m.htw <- var.test(DelPrete.orig.HTWM$`Anteroposterior diameter of the inlet`)

# HTB - Female
var.ap.f.htb <- var.test(DelPrete.orig.HTBF$`Anteroposterior diameter of the inlet`)

# HTB - Male
var.ap.m.htb <- var.test(DelPrete.orig.HTBM$`Anteroposterior diameter of the inlet`)

# TW - Female
var.ap.f.tw <- var.test(DelPrete.orig.TWF$`Anteroposterior diameter of the inlet`)

# TW -  Male
var.ap.m.tw <- var.test(DelPrete.orig.TWM$`Anteroposterior diameter of the inlet`)

# TB - Female
var.ap.f.tb <- var.test(DelPrete.orig.TBF$`Anteroposterior diameter of the inlet`)

# TB - Male
var.ap.m.tb <- var.test(DelPrete.orig.TBM$`Anteroposterior diameter of the inlet`)

# C - Female
var.ap.f.c <- var.test(DelPrete.orig.CF$`Anteroposterior diameter of the inlet`)

# C - Male
var.ap.m.c <- var.test(DelPrete.orig.CM$`Anteroposterior diameter of the inlet`)

# S - Female
var.ap.f.s <- var.test(DelPrete.orig.SF$`Anteroposterior diameter of the inlet`)

# S - Male
var.ap.m.s <- var.test(DelPrete.orig.SM$`Anteroposterior diameter of the inlet`)
```

## Pelvic Breadth Coefficients of Variation

```{r}
# HTW - Female
var.pb.f.htw <- var.test(DelPrete.orig.HTWF$`Pelvic breadth`)

# HTW - Male 
var.pb.m.htw <- var.test(DelPrete.orig.HTWM$`Pelvic breadth`)

# HTB - Female
var.pb.f.htb <- var.test(DelPrete.orig.HTBF$`Pelvic breadth`)

# HTB - Male
var.pb.m.htb <- var.test(DelPrete.orig.HTBM$`Pelvic breadth`)

# TW - Female
var.pb.f.tw <- var.test(DelPrete.orig.TWF$`Pelvic breadth`)

# TW - Male
var.pb.m.tw <- var.test(DelPrete.orig.TWM$`Pelvic breadth`)

# TB - Female
var.pb.f.tb <- var.test(DelPrete.orig.TBF$`Pelvic breadth`)

# TB - Male
var.pb.m.tb <- var.test(DelPrete.orig.TBM$`Pelvic breadth`)

# C - Female
var.pb.f.c <- var.test(DelPrete.orig.CF$`Pelvic breadth`)

# C - Male
var.pb.m.c <- var.test(DelPrete.orig.CM$`Pelvic breadth`)

# S - Female
var.pb.f.s <- var.test(DelPrete.orig.SF$`Pelvic breadth`)

# S - Male
var.pb.m.s <- var.test(DelPrete.orig.SM$`Pelvic breadth`)
```

## Bispinous Coefficients of Variation

```{r}
# HTW - Female
var.b.f.htw <- var.test(DelPrete.orig.HTWF$`Bi-spinous diameter`)

# HTW - Male
var.b.m.htw <- var.test(DelPrete.orig.HTWM$`Bi-spinous diameter`)

# HTB - Female
var.b.f.htb <- var.test(DelPrete.orig.HTBF$`Bi-spinous diameter`)

# HTB - Male
var.b.m.htb <- var.test(DelPrete.orig.HTBM$`Bi-spinous diameter`)

# TW - Female
var.b.f.tw <- var.test(DelPrete.orig.TWF$`Bi-spinous diameter`)

# TW - Male
var.b.m.tw <- var.test(DelPrete.orig.TWM$`Bi-spinous diameter`)

# TB - Female
var.b.f.tb <- var.test(DelPrete.orig.TBF$`Bi-spinous diameter`)

# TB - Male
var.b.m.tb <- var.test(DelPrete.orig.TBM$`Bi-spinous diameter`)

# C - Female
var.b.f.c <- var.test(DelPrete.orig.CF$`Bi-spinous diameter`)

# C - Male
c <- var.test(DelPrete.orig.CM$`Bi-spinous diameter`)

# S - Female
var.b.f.s <- var.test(DelPrete.orig.SF$`Bi-spinous diameter`)

# S - Male
var.b.m.s <- var.test(DelPrete.orig.SM$`Bi-spinous diameter`)
```

## AP Outlet Coefficients of Variation

```{r}
# HTW - Female
var.apo.f.htw <- var.test(DelPrete.orig.HTWF$`Anterioposterior space of outlet adj`)

# HTW - Male
var.apo.m.htw <- var.test(DelPrete.orig.HTWM$`Anterioposterior space of outlet adj`)

# HTB - Female
var.apo.f.htb <- var.test(DelPrete.orig.HTBF$`Anterioposterior space of outlet adj`)

# HTB - Male
var.apo.m.htb <- var.test(DelPrete.orig.HTBM$`Anterioposterior space of outlet adj`)

# TW - Female
var.apo.f.tw <- var.test(DelPrete.orig.TWF$`Anterioposterior space of outlet adj`)

# TW - Male
var.apo.m.tw <- var.test(DelPrete.orig.TWM$`Anterioposterior space of outlet adj`)

# TB - Female
var.apo.f.tb <- var.test(DelPrete.orig.TBF$`Anterioposterior space of outlet adj`)

# TB - Male
var.apo.m.tb <- var.test(DelPrete.orig.TBM$`Anterioposterior space of outlet adj`)

# C - Female
var.apo.f.c <- var.test(DelPrete.orig.CF$`Anterioposterior space of outlet adj`)

# C - Male
var.apo.m.c <- var.test(DelPrete.orig.CM$`Anterioposterior space of outlet adj`)

# S - Female
var.apo.f.s <- var.test(DelPrete.orig.SF$`Anterioposterior space of outlet adj`)

# S - Male
var.apo.m.s <- var.test(DelPrete.orig.SM$`Anterioposterior space of outlet adj`)
```

## TD Outlet Coefficients of Variation

```{r}
# HTW - Female
var.td.f.htw <- var.test(DelPrete.orig.HTWF$`Transverse diameter of the outlet`)

# HTW - Male
var.td.m.htw <- var.test(DelPrete.orig.HTWM$`Transverse diameter of the outlet`)

# HTB - Female
var.td.f.htb <- var.test(DelPrete.orig.HTBF$`Transverse diameter of the outlet`)

# HTB - Male
var.td.m.htb <- var.test(DelPrete.orig.HTBM$`Transverse diameter of the outlet`)

# TW - Female
var.td.f.tw <- var.test(DelPrete.orig.TWF$`Transverse diameter of the outlet`)

# TW - Male
var.td.m.tw <- var.test(DelPrete.orig.TWM$`Transverse diameter of the outlet`)

# TB - Female
var.td.f.tb <- var.test(DelPrete.orig.TBF$`Transverse diameter of the outlet`)

# TB - Male
var.td.m.tb <- var.test(DelPrete.orig.TBM$`Transverse diameter of the outlet`)

# C - Female
var.td.f.c <- var.test(DelPrete.orig.CF$`Transverse diameter of the outlet`)

# C - Male
var.td.m.c <- var.test(DelPrete.orig.CM$`Transverse diameter of the outlet`)

# S - Female
var.td.f.s <- var.test(DelPrete.orig.SF$`Transverse diameter of the outlet`)

# S - Male
var.td.m.s <- var.test(DelPrete.orig.SM$`Transverse diameter of the outlet`)
```

## Variability Matrices by Index Value
```{r}
# AP Inlet
var.mat.ap <- matrix(c(var.ap.f.htw,var.ap.f.htb,var.ap.f.tw, var.ap.f.tb, var.ap.f.c, var.ap.f.s, var.ap.m.htw, var.ap.m.htb, var.ap.m.tw, var.ap.m.tb, var.ap.m.c, var.ap.m.s),nrow=6, ncol=2)

# AP Outlet
var.mat.apo <- matrix(c(var.apo.f.htw,var.apo.f.htb,var.apo.f.tw, var.apo.f.tb, var.apo.f.c, var.apo.f.s, var.apo.m.htw, var.apo.m.htb, var.apo.m.tw, var.apo.m.tb, var.apo.m.c, var.apo.m.s),nrow=6, ncol=2)

# TD Outlet
var.mat.td <- matrix(c(var.td.f.htw,var.td.f.htb,var.td.f.tw, var.td.f.tb, var.td.f.c, var.td.f.s, var.td.m.htw, var.td.m.htb, var.td.m.tw, var.td.m.tb, var.td.m.c, var.td.m.s),nrow=6, ncol=2)

# Pelvic Breadth
var.mat.pb <- matrix(c(var.pb.f.htw,var.pb.f.htb,var.pb.f.tw, var.pb.f.tb, var.pb.f.c, var.pb.f.s, var.pb.m.htw, var.pb.m.htb, var.pb.m.tw, var.pb.m.tb, var.pb.m.c, var.pb.m.s),nrow=6, ncol=2)

# Bispinous
var.mat.b <- matrix(c(var.b.f.htw,var.b.f.htb,var.b.f.tw, var.b.f.tb, var.b.f.c, var.b.f.s, var.b.m.htw, var.b.m.htb, var.b.m.tw, var.b.m.tb, var.b.m.c, var.b.m.s),nrow=6, ncol=2)

```

## Var Matrices by Collection
```{r}
# HTW
htw.var.mat <- matrix(c(var.ap.f.htw, var.apo.f.htw, var.td.f.htw, var.pb.f.htw, var.b.f.htw, var.ap.m.htw, var.apo.m.htw, var.td.m.htw, var.pb.m.htw, var.b.m.htw), nrow=5)

# HTB
htb.var.mat <- matrix(c(var.ap.f.htb, var.apo.f.htb, var.td.f.htb, var.pb.f.htb, var.b.f.htb, var.ap.m.htb, var.apo.m.htb, var.td.m.htb, var.pb.m.htb, var.b.m.htb), nrow=5)

# TW
tw.var.mat <- matrix(c(var.ap.f.tw, var.apo.f.tw, var.td.f.tw, var.pb.f.tw, var.b.f.tw, var.ap.m.tw, var.apo.m.tw, var.td.m.tw, var.pb.m.tw, var.b.m.tw), nrow=5)

# TB
tb.var.mat <- matrix(c(var.ap.f.tb, var.apo.f.tb, var.td.f.tb, var.pb.f.tb, var.b.f.tb, var.ap.m.tb, var.apo.m.tb, var.td.m.tb, var.pb.m.tb, var.b.m.tb), nrow=5)

# C
c.var.mat <- matrix(c(var.ap.f.c, var.apo.f.c, var.td.f.c, var.pb.f.c, var.b.f.c, var.ap.m.c, var.apo.m.c, var.td.m.c, var.pb.m.c, var.b.m.c), nrow=5)

# S
s.var.mat <- matrix(c(var.ap.f.s, var.apo.f.s, var.td.f.s, var.pb.f.s, var.b.f.s, var.ap.m.s, var.apo.m.s, var.td.m.s, var.pb.m.s, var.b.m.s), nrow=5)
```


## Table and List of Variability
```{r}
# Matrix of all Coefficients of Variability
var.mat <- matrix(c(var.mat.ap, var.mat.apo, var.mat.td, var.mat.pb, var.mat.b), nrow=6)

# Vector of all female Cv
female.vals <- c(var.mat[1:6,1], var.mat[1:6,3], var.mat[1:6, 5], var.mat[1:6,7], var.mat[1:6,9])

# Vector of all male Cv
male.vals <- c(var.mat[1:6,2], var.mat[1:6,4], var.mat[1:6,6], var.mat[1:6,8], var.mat[1:6,10])

var.mat.final <- matrix(c(female.vals, male.vals), nrow=30)

# Tables of Matrices
ap.table <- as.table(var.mat.final[1:6,])
colnames(ap.table) = c("Female", "Male")
rownames(ap.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

apo.table <- as.table(var.mat.final[7:12,])
colnames(apo.table) = c("Female", "Male")
rownames(apo.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

td.table <- as.table(var.mat.final[13:18,])
colnames(td.table) = c("Female", "Male")
rownames(td.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

pb.table <- as.table(var.mat.final[19:24,])
colnames(pb.table) = c("Female", "Male")
rownames(pb.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

b.table <- as.table(var.mat.final[25:30,])
colnames(b.table) = c("Female", "Male")
rownames(b.table) = c("HTW", "HTB", "TW", "TB", "C", "S")

# List of Cvs
var.list <- list("AP Inlet" = ap.table,"AP Outlet" = apo.table, "TD Outlet" = td.table, "Pelvic Breadth" = pb.table, "Bispinous" = b.table)
```

## Significance of Variability

### AP Inlet
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.ap <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.HTB.ap <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.TW.ap <- with(DelPrete.orig.TW, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.TB.ap <- with(DelPrete.orig.TB, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.C.ap <- with(DelPrete.orig.C, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

sig.diff.S.ap <- with(DelPrete.orig.S, 
     asymptotic_test(`Anteroposterior diameter of the inlet`,Gender))

# Vector of Sig. Levels
var.sig.levels.ap <- c(sig.diff.HTW.ap$p_value, sig.diff.HTB.ap$p_value, sig.diff.TW.ap$p_value, sig.diff.TB.ap$p_value, sig.diff.C.ap$p_value, sig.diff.S.ap$p_value)

# AP Table w/ Sig Levels
ap.table.new <- as.table(matrix(c(var.mat.ap, var.sig.levels.ap), nrow=6))
colnames(ap.table.new) = c("Female", "Male", "Pvalue")
row.names(ap.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# AP Dataframe w/ Sig Levels
ap.mat <- matrix(c(var.mat.ap, var.sig.levels.ap), nrow=6)
ap.df <- as.data.frame(ap.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(ap.df)= c("Female", "Male", "P-value")
ap.df
```

### AP Outlet
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.apo <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.HTB.apo <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.TW.apo <- with(DelPrete.orig.TW, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.TB.apo <- with(DelPrete.orig.TB, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.C.apo <- with(DelPrete.orig.C, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

sig.diff.S.apo <- with(DelPrete.orig.S, 
     asymptotic_test(`Anterioposterior space of outlet adj`,Gender))

# Vector of Sig Levels
var.sig.levels.apo <- c(sig.diff.HTW.apo$p_value, sig.diff.HTB.apo$p_value, sig.diff.TW.apo$p_value, sig.diff.TB.apo$p_value, sig.diff.C.apo$p_value, sig.diff.S.apo$p_value)

# APO Table w/ Sig Levels
apo.table.new <- as.table(matrix(c(var.mat.apo, var.sig.levels.apo), nrow=6))
colnames(apo.table.new) = c("Female", "Male", "Pvalue")
row.names(apo.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# APO Dataframe w/ Sig Levels
apo.mat <- matrix(c(var.mat.apo, var.sig.levels.apo), nrow=6)
apo.df <- as.data.frame(apo.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(apo.df)= c("Female", "Male", "P-value")
apo.df
```

### TD Outlet
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.td <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.HTB.td <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.TW.td <- with(DelPrete.orig.TW, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.TB.td <- with(DelPrete.orig.TB, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.C.td <- with(DelPrete.orig.C, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

sig.diff.S.td <- with(DelPrete.orig.S, 
     asymptotic_test(`Transverse diameter of the outlet`,Gender))

# Vector of Sig Levels
var.sig.levels.td <- c(sig.diff.HTW.td$p_value, sig.diff.HTB.td$p_value, sig.diff.TW.td$p_value, sig.diff.TB.td$p_value, sig.diff.C.td$p_value, sig.diff.S.td$p_value)

# TD Table w/ Sig Levels
td.table.new <- as.table(matrix(c(var.mat.td, var.sig.levels.td), nrow=6))
colnames(td.table.new) = c("Female", "Male", "Pvalue")
row.names(td.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# TD Dataframe w/ Sig Levels
td.mat <- matrix(c(var.mat.td, var.sig.levels.td), nrow=6)
td.df <- as.data.frame(td.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(td.df)= c("Female", "Male", "P-value")
td.df
```

### Pelvic Breadth
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.pb <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.HTB.pb <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.TW.pb <- with(DelPrete.orig.TW, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.TB.pb <- with(DelPrete.orig.TB, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.C.pb <- with(DelPrete.orig.C, 
     asymptotic_test(`Pelvic breadth`,Gender))

sig.diff.S.pb <- with(DelPrete.orig.S, 
     asymptotic_test(`Pelvic breadth`,Gender))

# Vector of Sig Levels
var.sig.levels.pb <- c(sig.diff.HTW.pb$p_value, sig.diff.HTB.pb$p_value, sig.diff.TW.pb$p_value, sig.diff.TB.pb$p_value, sig.diff.C.pb$p_value, sig.diff.S.pb$p_value)

# PB Table w/ Sig Levels
pb.table.new <- as.table(matrix(c(var.mat.pb, var.sig.levels.pb), nrow=6))
colnames(pb.table.new) = c("Female", "Male", "Pvalue")
row.names(pb.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# PB Dataframe w/ Sig Levels
pb.mat <- matrix(c(var.mat.pb, var.sig.levels.pb), nrow=6)
pb.df <- as.data.frame(pb.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(pb.df)= c("Female", "Male", "P-value")
pb.df
```

### Bispinous
```{r}
# Asymptotic Test by Collection
sig.diff.HTW.b <- with(DelPrete.orig.HTW, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.HTB.b <- with(DelPrete.orig.HTB, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.TW.b <- with(DelPrete.orig.TW, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.TB.b <- with(DelPrete.orig.TB, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.C.b <- with(DelPrete.orig.C, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

sig.diff.S.b <- with(DelPrete.orig.S, 
     asymptotic_test(`Bi-spinous diameter`,Gender))

# Vector of Sig Levels
var.sig.levels.b <- c(sig.diff.HTW.b$p_value, sig.diff.HTB.b$p_value, sig.diff.TW.b$p_value, sig.diff.TB.b$p_value, sig.diff.C.b$p_value, sig.diff.S.b$p_value)

# B Table w/ Sig Levels
b.table.new <- as.table(matrix(c(var.mat.b, var.sig.levels.b), nrow=6))
colnames(b.table.new) = c("Female", "Male", "Pvalue")
row.names(b.table.new) = c("HTW", "HTB", "TW", "TB", "C", "S")

# B Dataframe w/ Sig Levels
b.mat <- matrix(c(var.mat.b, var.sig.levels.b), nrow=6)
b.df <- as.data.frame(b.mat, row.names = c("HTW","HTB", "TW", "TB", "C", "S"))
colnames(b.df)= c("Female", "Male", "P-value")
b.df
```

## New Variations List
```{r}
# List of Cvs w/ Sig Levels
var.list.new <- list("AP Inlet" = ap.table.new,"AP Outlet" = apo.table.new, "TD Outlet" = td.table.new, "Pelvic Breadth" = pb.table.new, "Bispinous" = b.table.new)

# List of Cv Dataframes w/ Sig Levels
final.var.list <- list("AP Inlet"=ap.df, "AP Outlet"=apo.df, "TD Outlet"=td.df, "Pelvic Breadth"=pb.df, "Bispinous"=b.df)
final.var.list
```
Using 0.05 conf. level: C - AP Outlet, C - Pelvic Breadth have significant difference
Using 0.1 conf. level: HTB, TW, C, S - AP Outlet, HTB, C - Pelvic Breadth 
Very small amount of collections by index value have a significant difference between males and females
